### 2020-04-11 paper/paper.tex
intro and overview

%% This is an abbreviated template from http://www.sigplan.org/Resources/Author/.

\documentclass[acmsmall,review,authorversion]{acmart}
\acmDOI{}
% \acmJournal{FACMP}
\acmVolume{CSCI 5535}
\acmNumber{Spring 2020}

\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Name of the Title is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{TBD}
\email{tbd@colorado.edu}
\author{TBD}
\email{tbd@colorado.edu}
\affiliation{%
  \institution{University of Colorado Boulder}
}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  TBD
\end{abstract}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

One of the more powerful tools in use by functional programmers are engines for
searching over code bases given the type signature of the desired function. Such
tools provide the entry-point for practical code reuse and management of large
code bases. Tools such as Hoogle(for Haskell) provide a best-in-class search
functionality to polymorphic functional languages, but the same tools are not as
prevalent for dependently-typed languages.

There are a number of factors that have prevented robust search features from
entering the dependently-typed landscape. The primary setback is the
undecidability of type equality(isomorphism) in the presence of dependent types.
For search over types to be performed there must be some way of saying whether
two types are equal or not. This introduces many of the nuances of equality into
the confusion of trying to build a practical tool.

We attempt to solve this problem by defining an abstraction which defines a
search engine for dependent types, that can scale and adapt to appropriate
definitions of type equality. It is reasonable for a user to want some control
over how weak/strong the equalities for search are. Some may want strong
definitional equalities, others may accept the more computationally burdensome(and
semi-decidable) isomorphisms. We give an approach that leaves this control to
the user, and also leaves open the possibility for very powerful forms of
program search.

Another deficiency of current search engines for program source is the lack of
being able to specify complex functional properties over the domain of search.
Input-Output examples and algebraic properties offer two great sources of
specification for search, yet they are not used, typically because of the lack
of some method for verifying these properties. Fortunately, a dependently
typed language offers a natural way of handling these properties.

Practically, we provide all of the abstractions formalized in Agda and presented
as concrete tool for search over functions with input-output examples. The
implementation is capable of being extended to support more expressive
properties which in turn will provide more power to a user to specify complex
queries.

\section{Overview}

Let's start with an example of a search query over a non-dependent type, given
as a type:

$$
\text{query} : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbb{N}
$$

Solutions for this query would obviously include the usual arithmetic
operations. But let's say that we wanted to further specify the query by
providing some input-output examples:

$$
\begin{array}{ccc}
  \text{input 1} & \text{input 2} & \text{output}\\
  0 & 0 & 0\\
  2 & 2 & 4\\
  1 & 2 & 3\\
  ... &...&...
\end{array}
$$

We could devise the obvious procedure of checking the resulting functions
against these examples and discarding those that are not consistent. In fact,
this is exactly what we will do. But to jump to the algorithm would fail to
realize the deeper type-theoretic possibilities here. See, the examples above
can be encoded quite easily as a type for our query. The type below is
consistent with the query:

$$
\begin{array}{c}
\Sigma[\text{f} \in \mathbb{N} \rightarrow \mathbb{N} \rightarrow
  \mathbb{N}] ((f\ 0\ 0\ \equiv 0) \times (f\ 2\ 2\ \equiv 4) \times (f\ 1\ 2\
  \equiv 3))
\end{array}
$$

This is a \textit{dependent product} type that encodes our query as searching
for a term that satisfies the type we gave along with proofs that the provided
function produces certain results under evaluation.

\section{(Contribution 1)}

\section{(Contribution 2)}

\section{Empirical Evaluation}

\section{Related Work}

\section{Conclusion}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
TBD
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}
\end{document}