### 2020-04-25 paper/paper.tex
technical sections

%% This is an abbreviated template from http://www.sigplan.org/Resources/Author/.

\documentclass[acmsmall,review,authorversion]{acmart}
\acmDOI{}
% \acmJournal{FACMP}
\acmVolume{CSCI 5535}
\acmNumber{Spring 2020}

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
% \newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
% \newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
% \newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}

\newcommand{\NC}{(\textbf{Needs Citation})}

\newcommand{\term}[1]{\textsf{#1}}
\newcommand{\type}[1]{\textsf{#1}}

\newcommand{\Carrier}{\term{Carrier}}
\newcommand{\Universe}{\mathcal{U}}
\newcommand{\Set}{\type{Set}}
\newcommand{\Type}{\type{Type}}
\newcommand{\Boolean}{\term{Boolean}}
\newcommand{\true}{\term{true}}
\newcommand{\false}{\term{false}}
\newcommand{\Number}{\term{Number}}
\newcommand{\Interp}[1]{\llbracket #1 \rrbracket}
\newcommand{\?}{\stackrel{?}{\approx}}
\newcommand{\search}{\term{search}}

\newcommand{\data}[2]{\textsf{data } #1 : #2 \textsf{ where}}
\newcommand{\record}[2]{\textsf{record } #1 : #2 \textsf{ where}}

\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Embedded Program Search in Agda}

%%
%% The "author" command and its associa

%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Jack Martin}
\email{John.P.Martin@colorado.edu}
\author{Michael Dresser}
\email{tbd@colorado.edu}
\affiliation{%
  \institution{University of Colorado Boulder}
}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
The problem of Program Search - looking for a function which satisfies a query -
is one of the more practical tools a developer can use when programming. Search
tools such as Hoogle provide an entrypoint for a language eco-system, but
unfortunately often don't accept query information other than types. Here we
present a search tool for searching over types and examples as well as a general
framework for doing program search over dependent types. We also provide a
type-theoretic interpretation of the program search problem as a special case of
the more general synthesis problem and discuss how this presents a tool for
furthering the solution to both problems.
\end{abstract}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

One of the more powerful tools in use by functional programmers are engines for
searching over code bases given the type signature of the desired function. Such
tools provide the entry-point for practical code reuse and management of large
code bases. Tools such as Hoogle(for Haskell) provide a best-in-class search
functionality to polymorphic functional languages, but the same tools are not as
prevalent for dependently-typed languages.

There are a number of factors that have prevented robust search features from
entering the dependently-typed landscape. The primary setback is the
undecidability of type equality(isomorphism) in the presence of dependent types.
For search over types to be performed there must be some way of saying whether
two types are equal or not. This introduces many of the nuances of equality into
the confusion of trying to build a practical tool.

We attempt to solve this problem by defining an abstraction which defines a
search engine for dependent types, that can scale and adapt to appropriate
definitions of type equality. It is reasonable for a user to want some control
over how weak/strong the equalities for search are. Some may want strong
definitional equalities, others may accept the more computationally burdensome(and
semi-decidable) isomorphisms. We give an approach that leaves this control to
the user, and also leaves open the possibility for very powerful forms of
program search.

Another deficiency of current search engines for program source is the lack of
being able to specify complex functional properties over the domain of search.
Input-Output examples and algebraic properties offer two great sources of
specification for search, yet they are not used, typically because of the lack
of some method for verifying these properties. Fortunately, a dependently
typed language offers a natural way of handling these properties.

Practically, we provide all of the abstractions formalized in Agda and presented
as concrete tool for search over functions with input-output examples. The
implementation is capable of being extended to support more expressive
properties which in turn will provide more power to a user to specify complex
queries.

\section{Overview}

Let's start with an example of a search query over a non-dependent type, given
as a type:

$$
\text{query} : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbb{N}
$$

Solutions for this query would obviously include the usual arithmetic
operations. But let's say that we wanted to further specify the query by
providing some input-output examples:

$$
\begin{array}{ccc}
  \text{input 1} & \text{input 2} & \text{output}\\
  0 & 0 & 0\\
  2 & 2 & 4\\
  1 & 2 & 3\\
  ... &...&...
\end{array}
$$

We could devise the obvious procedure of checking the resulting functions
against these examples and discarding those that are not consistent. In fact,
this is exactly what we will do. But to jump to the algorithm would fail to
realize the deeper type-theoretic possibilities here. See, the examples above
can be encoded quite easily as a type for our query. The type below is
consistent with the query:

$$
\begin{array}{c}
\Sigma[\text{f} \in \mathbb{N} \rightarrow \mathbb{N} \rightarrow
  \mathbb{N}] ((f\ 0\ 0\ \equiv 0) \times (f\ 2\ 2\ \equiv 4) \times (f\ 1\ 2\
  \equiv 3))
\end{array}
$$

This is a \textit{dependent product} type that encodes our query as searching
for a term that satisfies the type we gave along with proofs that the provided
function produces certain results under evaluation.

\section{Preliminaries}

\todo[inline]{Note for Benno and Evan: The citations are absent for now. We will
add these for the final paper but decided to leave them out for now until the
structure of the paper is fixed. The notes for ``Needs citation'' indicate
positions where we have a reference saved somewhere. It is not an indication
that we need to still find it. }

\subsection{Universes} One of the advantages of dependent type theories is their
ability to formalize their own type system and meta-theories within
themselves(\textbf{Needs citation}). These techniques typically use a structure
called a \textit{Universe}, which is defined as a type along with an
interpretation function which takes an element of that type and computes some
type in the type system. Formally:

$$
\begin{array}{l}
\record{\Universe}{\Set}\\
\quad  \begin{array}{rl}
         \Carrier &: \Set \\
         \Interp{\_} &: \Carrier \rightarrow \Set
       \end{array}
\end{array}
$$

An example of such a universe containing booleans, natural numbers, and
functions on either is given here:

$$
\begin{array}{cc}
\begin{array}{l}
\data{\Type}{\Set}\\
\quad \begin{array}{rl}
        \textsf{Number} :& \Type\\
        \textsf{Boolean} :& \Type\\
        \_\Rightarrow\_ :& \Type \rightarrow \Type \rightarrow \Type
      \end{array}
\end{array}
&
\begin{array}{l}
  \Interp{\_}_{\Type} : \Type \rightarrow \Set\\
  \Interp{\textsf{Number}}_{\Type}     = \mathbb{N}\\
  \Interp{\textsf{Boolean}}_{\Type}    = \mathbb{B}\\
  \Interp{t_1 \Rightarrow t_2}_{\Type} \ \, = \Interp{t_1} \rightarrow \Interp{t_2}
\end{array}
\end{array}
$$

Soon, we will define a universe of types for use in program search. We will
spend most of the rest of the paper defining a suitable $\Carrier$ type that
respects the various complexities that arise from the presence of dependent
types.

\subsection{Simulating First Class Modules}

We will need a coarse simulation of Agda's module structure to store entries for
search. This will be accomplished by collecting elements of a $\Carrier$ type
along with the terms they encode into a homogeneous list. The first element will
be the coded type, and the second element(which depends on the first), will be
the term of the type that is interpreted from the code:

$$
\textsf{Module} = \Sigma[ \tau \in \Carrier ] \Interp{ \tau }
$$

Carrying on with our example universe above, an example module containing some
common functions and elements:

$$
\begin{array}{rl}
  \textsf{example} =& (\Boolean\ ,\ \true)\\
                  ::& (\Number\ ,\ 5 )\\
                  ::& (\Boolean\ ,\ \false )\\
                  ::& (\Number \Rightarrow \Boolean\ ,\ (== 5) )\\
                  ::&[]
\end{array}
$$

\section{Searching over Universe Modules}

To support basic type search over these modules we will first have to discuss
what it means for a query type to ``match'' a type in a collection. This clearly
introduces the need for a notion of equality over types, which complicates
matters quickly. Which form of equality of types is appropriate for use in
program search? Fortunately, there has been some research done on this problem
and the current accepted equality is isomorphism of types\NC. This
will equate curried function types to their uncurried counterparts for
example(i.e. $(A \times B) \rightarrow C \approx A \rightarrow B \rightarrow
C$).

A problem with using isomorphism over dependent types is that it is a generally
undecidable problem\NC. The field is currently studying certain decidable
subsets of the problem thus we opt to maximize flexibility by parameterizing our
methods over a provided decidable type equality. The problem of type-directed
program search then just becomes the act of running the decision procedure for
this equality over the coded-types in our module. Thus, a searchable universe
can be defined:
\todo{define isEquiv}
$$
\begin{array}{l}
  \record{\Universe_S}{\Set}\\
  \quad \begin{array}{ll}
          U &: \Universe\\
         \_\approx\_ &: (a\ b\ : \Carrier\ U) \rightarrow \Set\\
         \approx-\textsf{isEquiv} &: \textsf{IsEquivalence } (\Carrier\ U)\ \_\approx\_\\
         \_\?\_ &: \forall (a\ b : \Carrier\ U) \rightarrow \textsf{Dec } (a \approx b)
        \end{array}
\end{array}
$$

Now we are ready to define an abstract search procedure for modules of a
decidable universe. The type of the search function is:

$$
\begin{array}{l}
  \search : (u : \Universe_S) \rightarrow (m : \textsf{Module}_u)\rightarrow (c : \Carrier\ (U\ u))
  \rightarrow \textsf{Maybe }\Interp{c}
\end{array}
$$

The procedure takes in a decidable universe $u$, a module of that universe $m$,
and a coded-type $c$. The output type of the function is possibly an inhabitant
of the interpretation of $c$ provided that there is in fact one in the module.

\section{A Universe for Searching over Dependent Types}

Thus far, the formalism we have given should allow any non-dependent program
to be searched. To add support for programs with type dependencies we will need
to slightly alter our definition of $\textsf{Module}$. The main idea is that the
module will need to be simultaneously defined with the interpretation function
so that types defined in a ``cons'' of a module can depend on the terms found in
earlier entries.

The definition of this universe is given in Figure \ref{fig1}.
\todo[inline]{Note for Benno and Evan: There is not any explanation or elaboration for this
  part of the formalism yet. There are also a few missing constructors. We were
  stuck on this for the better part of the last week and just figured out how to
formalize it last night and it isn't quite finished. In the final paper this
will be much, much more spelled out over 3-5 pages.}

\begin{figure}[t]
  $$
  \begin{array}{cc}
  \begin{array}{l}
    \data{\textsf{Module}}{\mathbb{N} \rightarrow \Set}\\
    \quad \begin{array}{ll}
            []        &: \textsf{Module } 0\\
            \_::\_,\_ &: \forall \{ n \}\\
                      &\rightarrow  (\Gamma : \textsf{Module } n)\\
                      &\rightarrow  (t : \textsf{TypeCode } n)\\
                      &\rightarrow  \Interp{ t }_\Gamma \\
                      &\rightarrow \textsf{Module (suc }  n)
          \end{array}
  \end{array}
                        &
  \begin{array}{l}
    \data{\textsf{TypeCode}}{\mathbb{N} \rightarrow \Set}\\
    \quad
    \begin{array}{ll}
      U &: \forall \{ n \} \rightarrow \textsf{TypeCode } n\\
      \# &: \forall \{ n\ m \} \rightarrow n < m \rightarrow \textsf{TypeCode } m\\
      \_\Rightarrow\_ &: \forall \{ n \}\ (t_1\ t_2 : \textsf{TypeCode } n)
                        \rightarrow \textsf{TypeCode } n
    \end{array}

  \end{array}
  \end{array}
  $$
  $$
  \begin{array}{l}
    \Interp{\_}\_ : \forall \{ n \} \rightarrow \textsf{TypeCode } n \rightarrow
    \textsf{Module } n \rightarrow \Set\\
    \Interp{ U }_\Gamma = \Set\\
    \Interp{ t_1 \Rightarrow t_2 }_\Gamma = \Interp{t_1}_\Gamma \rightarrow \Interp{t_2}_\Gamma\\
    \Interp{ \#\ \textsf{suc-leq-suc ( zero-leq-n)}}_{(\Gamma :: U ,
    \textsf{term})} = \textsf{term}\\
    \Interp{ \#\ \textsf{suc-leq-suc ( zero-leq-n)}}_{(\Gamma :: t ,
    \textsf{term})} = \langle\textsf{Error Case - Impossible under assumptions}\rangle\\
    \Interp{ \#\ \textsf{suc-leq-suc ( suc-leq-suc } n )}_{(\Gamma :: t ,
    \textsf{term})} = \Interp{ \#\ \textsf{suc-leq-suc } n }_\Gamma
  \end{array}
  $$
  \caption{The Universe for Search over Dependent Types}
  \label{fig1}
\end{figure}

\section{Empirical Evaluation}

\todo[inline]{We were unsure what to include in this section because most of our
  project was correct by construction. Benno and Evan, we will touch base with
  you about your thoughts as to what to include here. We brainstormed a little
  bit but couldn't settle on anything that made too much sense. Hoping that the
  sections above will be suitable for this week's version of the draft. One idea
  we had would be to show off an example of why the embedded program search is
  so powerful/useful. This seems to be a somewhat empirical claim that we could
  spend some time fleshing out.}

\section{Related Work}

\todo[inline]{Edit: Make changes as Benno suggested}

\paragraph{Program Search} Program Search, or the search of functions in some
library, has led to such tools as Hoogle, and is the main category that this
work could be included under. In the past, none of these tools use any
functional properties or examples to help guide the search. Ours is the first
tool to provide this functionality.

\textit{draft note: make sure to look into the search programs for other languages.}

\paragraph{Program Synthesis} As was mentioned in Section \# above, program
search can be viewed as a special case of the more general program synthesis
problem. It seems hopeful that program search can be later used as a tool in
program synthesis by treating it as an atomic operation for larger synthesis
procedures.

\paragraph{Type Isomorphism} Isomorphisms over types acts as the equivalence
relation that we use for search in our approach. In reality our approach defines
a slightly more robust isomorphism relation over types \textit{and} examples.
This is slightly more powerful than just the types themselves. In future work we
will look at the theoretical basis for this.

\section{Conclusion}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
TBD
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}
\end{document}